---
title: "FINAL"
author: "Ahmed Adeeb AbdulCader Mohammed Shaheed"
date: "2022-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Adeeb/Documents/Final Exam")
library(tidyverse)
library(dplyr)
```

## R Markdown

#PART A

#Question1a
```{r include=TRUE}

popAL<-read.csv("sub-est2016_1.csv",head=T)
popAK<-read.csv("sub-est2016_2.csv",head=T)
popCombined <- merge(popAL,popAK,by=c('NAME','STNAME'),all.x=T)
```
```{r include=TRUE}

popUS<-read.csv("sub-est2016_all.csv",head=T)

unique(popUS$STNAME)
```
```{r include=TRUE}

popUSstates = subset(popUS, NAME == STNAME,select=c('NAME','STNAME','POPESTIMATE2012'))
```
```{r include=TRUE}
popUSstates2 = subset(popUSstates, STNAME == NAME,select=c('NAME','STNAME','POPESTIMATE2012'))
```
```{r include=TRUE}
areaUS <- read.csv('https://raw.githubusercontent.com/jakevdp/PythonDataScienceHandbook/master/notebooks/data/state-areas.csv')

popArea <- left_join(popUSstates, areaUS, by=c('STNAME'='state'))
```
```{r include=TRUE}

popArea$popdensity <- (popArea$POPESTIMATE2012/popArea$area..sq..mi.)
```
```{r include=TRUE}

ecnUS <-read.csv("ECN_2012_US_52A1.csv",head=T)
ecnUS <- ecnUS[-1,]
```
```{r include=TRUE}

ecnUS$RCPTOT <- as.numeric(RCPTOT)

rev <-ecnUS%>%
  group_by(GEO.display.label,NAICS.display.label)%>%
  summarise(RCPTOT=sum(RCPTOT))
```
```{r include=TRUE}
  
rev <- na.omit(rev)

popRev <- left_join(popArea, rev, by=c('STNAME'='GEO.display.label'))

popRev2 <- popRev%>%
  group_by(STNAME,popdensity)%>%
  summarise(RCPTOT=sum(RCPTOT))
```
```{r include=TRUE}

ggplot(data = df11, mapping = (aes(x = `popdensity`, y = `RCPTOT`)))+
  geom_point()+
  geom_smooth(method = lm, se = FALSE)
```
```{r include=TRUE}

load("~/Desktop/Class/Data 210/nes.rda")

table(nes$interest_voted2008)
```{r include=TRUE}

nes$interest_voted2008[nes$interest_voted2008 == 1] <- "Voted"
nes$interest_voted2008[nes$interest_voted2008 == 2] <- "Didn't Vote"
nes$interest_voted2008[nes$interest_voted2008 == -8] <- "Don't Know"
nes$interest_voted2008[nes$interest_voted2008 == -9] <- "NA"
table(nes$interest_voted2008)
```
```{r include=TRUE}

table(nes$interest_whovote2008)
```
```{r include=TRUE}

nes$interest_whovote2008[nes$interest_whovote2008 == 1] <- "Barack Obama"
nes$interest_whovote2008[nes$interest_whovote2008 == 2] <- "John McCain"
nes$interest_whovote2008[nes$interest_whovote2008 == 5] <- "Other"
nes$interest_whovote2008[nes$interest_whovote2008 == -8] <- "Don't Know"
nes$interest_whovote2008[nes$interest_whovote2008 == -1 | 
                           nes$interest_whovote2008 == -9] <- "NA"
table(nes$interest_whovote2008)
```
```{r include=TRUE}

with(nes[nes$interest_voted2008 == "Voted",],
     wpct(interest_whovote2008, weight_full)) * 100
```
```{r include=TRUE}

table(nes$ftgr_fedgov)
```
```{r include=TRUE}


nes <- nes %>% filter(ftgr_fedgov >= 0 & ftgr_fedgov <= 100)

table(nes$ftgr_fedgov)
```
```{r include=TRUE}


weighted.mean(nes$ftgr_fedgov, w = nes$weight_full)
```
```{r include=TRUE}


table(nes$prevote_regpty)
```
```{r include=TRUE}

nes$prevote_regpty[nes$prevote_regpty == 1] <- "Democrat"
nes$prevote_regpty[nes$prevote_regpty == 2] <- "Republican"
nes$prevote_regpty[nes$prevote_regpty == 4 | 
                     nes$prevote_regpty == 5 | 
                     nes$prevote_regpty == -1 | 
                     nes$prevote_regpty == -8 | 
                     nes$prevote_regpty == -9] <- "NA"
table(nes$prevote_regpty)
```
```{r include=TRUE}


dems <- nes %>% filter(prevote_regpty == "Democrat")

reps <- nes %>% filter(prevote_regpty == "Republican")

weighted.mean(dems$ftgr_fedgov, w = dems$weight_full) -
weighted.mean(reps$ftgr_fedgov, w = reps$weight_full)
```