---
title: "HW WEEK 7"
author: "Ahmed Adeeb AbdulCader Mohammed Shaheed"
date: "2022-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## R Markdown

#Question 1a

#Felony voting patterns would be hard to decipher as we aren't able to create a set of random study participants with varying levels of felony convictions. It is also highly improbable that anyone asked what they would do if they were/weren't a felon could objectively answer the question without any bias. 

#Instead, the best option we have to make sense of this data is to build a study with participants who have very similar backgrounds and only differ in their status of being felons or not, and then build a study to compare their behavior and check for any correlations.

#Question 1b

#The authors are interested in studying the causal effect of an encouragement to register and to vote on released felons.

#Two treatment groups and one control group.Each member of the two groups was sent a treatment letter in addition to their official mail-in ballot. The treatment letters were similar except the second group's treatment letter provides reassurance that they wouldn't be asked about their prior status as a felon.

#The authors used block randomization, with blocks for individual crime types.


#Question 1c

```{r include=TRUE}

load("~/felons.RData")
head(felons)
```
```{r include=TRUE}

table(felons$returntoprison)
```
```{r include=TRUE}

felons <- felons[!felons$returntoprison == 1,]

felons$treatment_collapsed <- FALSE
felons$treatment_collapsed[felons$treatment == 2 | felons$treatment == 3] <- TRUE

table(felons$treatment_collapsed)
```
#Question 1d

```{r include=TRUE}

age.test <- t.test(felons$age ~ felons$treatment_collapsed)

time.served.test <- t.test(felons$days_served ~ felons$treatment_collapsed)

years.released.test <- t.test(felons$yrs_since_release ~ felons$treatment_collapsed)

vote08.test <- t.test(felons$vote08 ~ felons$treatment_collapsed)

t.tests <- data.frame(Variable = character(0),
                      Estimate = numeric(0),
                      P.Value = numeric(0))

t.tests <- rbind(t.tests, c("Age", age.test$estimate, age.test$p.value))
t.tests <- rbind(t.tests, c("Time Served", time.served.test$estimate, 
                            time.served.test$p.value))
t.tests <- rbind(t.tests, c("Years Released", years.released.test$estimate, 
                            years.released.test$p.value))
t.tests <- rbind(t.tests, c("Vote 08", vote08.test$estimate, vote08.test$p.value))

colnames(t.tests) <- c("Variable", "Control Estimate", "Treated Estimate", "P.Value")


t.tests
```
```{r include=TRUE}

summary(lm(treatment_collapsed ~ felony_type, data = felons))
```
```{r include=TRUE}

summary(lm(treatment_collapsed ~ age + days_served + yrs_since_release + 
             vote08 + felony_type, data = felons))
```

#Question 1e

#The likelihood that members of the untreated group would register to vote was approximately.058 or 5.8%, but the likelihood that members of the treated group would do so was approximately.017 higher, at.075 or 7.5%.
#The likelihood of the untreated group voting was approximately.029 or 2.9%, whereas the likelihood of the treated group voting was approximately.009 greater than that, at .038 or 3.8%.
```{r include=TRUE}

summary(lm(registered ~ treatment_collapsed, data = felons))
```
```{r include=TRUE}

summary(lm(vote12 ~ treatment_collapsed, data = felons))
```
```{r include=TRUE}

summary(lm(registered ~ treatment_collapsed + age + days_served + 
             yrs_since_release + vote08 + felony_type, data = felons))

summary(lm(vote12 ~ treatment_collapsed + age + days_served + 
             yrs_since_release + vote08 + felony_type, data = felons))
```