---
title: "AbdulCader_Mohammed_Shaheed"
author: "Ahmed Adeeb AbdulCader Mohammed Shaheed"
date: "2022-09-26"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
library(tidyverse) 
library(dslabs)
knitr::opts_chunk$set(echo = TRUE)
```
## R Markdown

#-------------------------------------------------
# Topic 6 HW
# Name: Ahmed Adeeb AbdulCader Mohammed Shaheed
#-------------------------------------------------

#-------------------------------------------------
#Question-1

```{r include=TRUE}
data(gapminder)

```
#Question-2

```{r include=TRUE}
class(gapminder)

```
#Question-3
```{r include=TRUE}
dim(gapminder)

```

#Question-4
```{r include=TRUE}
names(gapminder)

```

#Question-5
```{r include=TRUE}
attributes(gapminder$life_expectancy)
attributes(gapminder$fertility)
```
#Question-6
```{r include=TRUE}
#It wasn't included by creator of dataset
```
#Question-7
```{r include=TRUE}
View(gapminder)

gapminder$population[gapminder$country == "Norway" & gapminder$year == 1962]
```

#Question-8

```{r include=TRUE}
unique(gapminder$country[gapminder$life_expectancy < 40])

```
```{r include=TRUE}
gapminder$country[gapminder$life_expectancy < 40 & gapminder$year > 2000]

```
#Problem2
#PartA
#Question-1

```{r include=TRUE}
recent_grads <- read_csv("Newsweek_graduation_subset.csv")
```

#Question-2

```{r include=TRUE}
dim(recent_grads)
```

```{r include=TRUE}

class(recent_grads)
```

```{r include=TRUE}

colnames(recent_grads)
```

#Question-3
#column attributes such as row and column names are included along with column specifications. there isnt any hidden attributes for columns because we imported a csv file.
```{r include=TRUE}
attributes(recent_grads)
```
#Part B
#Question-1
```{r include=TRUE}
recent_grads$Major_category[which.max(recent_grads$Median)]
```
```{r include=TRUE}
recent_grads$Major_category[which.min(recent_grads$Median)]
```
#Question-2
```{r include=TRUE}
mean(recent_grads$Median[recent_grads$Major_category == "Engineering"])
```
```{r include=TRUE}
mean(recent_grads$Median[recent_grads$Major_category == "Education"])
```


#Question-4

```{r include=TRUE}
unique(recent_grads$Major_category)
```
```{r include=TRUE}
mean(recent_grads$ShareWomen[recent_grads$Major_category == "Engineering"])
```
```{r include=TRUE}
mean(recent_grads$ShareWomen[recent_grads$Major_category == "Physical Sciences"])
```
```{r include=TRUE}
mean(recent_grads$ShareWomen[recent_grads$Major_category == "Computers & Mathematics"])
```
```{r include=TRUE}
mean(recent_grads$ShareWomen[recent_grads$Major_category == "Biology & Life Science"])
```


