---
title: "AbdulCader_Mohammed_Shaheed"
author: "Ahmed Adeeb AbdulCader Mohammed Shaheed"
date: "2022-09-26"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
library(tidyverse) 
library(dslabs)
knitr::opts_chunk$set(echo = TRUE)
```
## R Markdown

#-------------------------------------------------
# Topic 6 HW
# Name: Ahmed Adeeb AbdulCader Mohammed Shaheed
#-------------------------------------------------

#-------------------------------------------------
#Question-1

```{r include=TRUE}
data(gapminder)

```
#Question-2

```{r include=TRUE}
class(gapminder)

```
#Question-3
```{r include=TRUE}
dim(gapminder)

```

#Question-4
```{r include=TRUE}
names(gapminder)

```

#Question-5
```{r include=TRUE}
attributes(gapminder$life_expectancy)
attributes(gapminder$fertility)
```
#Question-6
```{r include=TRUE}
#It wasn't included by creator of dataset
```
#Question-7
```{r include=TRUE}
View(gapminder)

gapminder$population[gapminder$country == "Norway" & gapminder$year == 1962]
```

#Question-8

```{r include=TRUE}
unique(gapminder$country[gapminder$life_expectancy < 40])

```
```{r include=TRUE}
gapminder$country[gapminder$life_expectancy < 40 & gapminder$year > 2000]

```
#Problem2
#PartA
#Question-1

```{r include=TRUE}
recent_grads <- read_csv("Newsweek_graduation_subset.csv")
```

#Question-2

```{r include=TRUE}
dim(recent_grads)
```

```{r include=TRUE}

class(recent_grads)
```

```{r include=TRUE}
names(recent_grads)

colnames(recent_grads)
```

```{r include=TRUE}

view(recent_grads)
```
#column attributes such as row and column names are included along with column specifications. there isnt any hidden attributes for columns because we imported a csv file.
```{r include=TRUE}
attributes(recent_grads)
```
#Part B
#Question-1
```{r include=TRUE}
which.max(recent_grads$Median)
recent_grads[1,]
```
```{r include=TRUE}
which.min(recent_grads$Median)
recent_grads[173,]
```
#Question-2
##(a) Find major category with highest and lowest median income 

```{r include=TRUE}
mean(recent_grads$Median[recent_grads$Major_category == "Engineering"])
```
```{r include=TRUE}
mean(recent_grads$Median[recent_grads$Major_category == "Education"])
```


##(b) Are there any professions with the top 10 highest median income and highest share of women? 

```{r include=TRUE}
top.income <- recent_grads[1:10,]
```
```{r include=TRUE}
high.women.share<-recent_grads%>%
  arrange(desc(ShareWomen))%>%
  head(10)
```
# (c) Are women underrepresented in STEM fields still? 

```{r include=TRUE}
mean(recent_grads$ShareWomen, na.rm=TRUE)
```
```{r include=TRUE}
stem<-subset(recent_grads[(recent_grads$Major_category=="Engineering")
                          |(recent_grads$Major_category=="Physical Sciences")
                          |(recent_grads$Major_category=="Biology & Life Science")
                          |(recent_grads$Major_category=="Computers & Mathematics"),])
```


