---
title: "DATA 310"
author: "Ahmed Adeeb AbdulCader Mohammed Shaheed"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(scales)
library(tigerstats)


knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#-------------------------------------------------
# Homework 2
# Name: Ahmed Adeeb AbdulCader Mohammed Shaheed
#-------------------------------------------------

#-------------------------------------------------

#Question-1

You may determine the possibility of receiving a certain value for a variable, or a range of values, using a probability distribution. It may be used to calculate the likelihood of achieving a specific result. You might predict your likelihood of rolling two sixes in advance, for instance, if you're rolling two dice.



#Question-2

#a

```{r include=TRUE}
(pnorm(30, mean = 26, sd = sqrt(54)) - pnorm(24, mean = 26, sd = sqrt(54))) * 100

pnormGC(c(24, 30), region = "between", mean = 26, sd = sqrt(54), graph = TRUE)

```

#b
```{r include=TRUE}
((100 - 95) / 2 + 95) / 100

```
```{r include=TRUE}
qnorm(0.975, 26, sqrt(54))

```
```{r include=TRUE}
qnorm((1-0.975), 26, sqrt(54))

```
```{r include=TRUE}
pnormGC(c(11.59726, 40.40274), region = "between", mean = 26, sd = sqrt(54), graph = TRUE)

```
#Question-3

#a

```{r include=TRUE}
(pnorm(50, mean = 43, sd = sqrt(23)) - pnorm(40, mean = 43, sd = sqrt(23))) * 100

```

```{r include=TRUE}
pnormGC(c(40, 50), region = "between", mean = 43, sd = sqrt(23), graph = TRUE)

```

#b
```{r include=TRUE}
((100 - 99) / 2 + 99) / 100

```
```{r include=TRUE}
qnorm(0.995, mean = 43, sd = sqrt(23))

```

```{r include=TRUE}
qnorm((1-0.995), mean = 43, sd = sqrt(23))


```

```{r include=TRUE}
pnormGC(c(30.64676, 55.35324), region = "between", mean = 43, sd = sqrt(23), graph = TRUE)

```
#Question-4

#a
```{r include=TRUE}

numdice = 4

numsims = 1000000

maxnumber <- matrix(NA, nrow = numsims, ncol = 1)  

set.seed(12221979)

for (s in 1:numsims) {
  
  uniform <- runif(n = numdice, min = 0, max = 1)

  dice <- matrix(-9, nrow = numdice, ncol = 1)  
  dice[uniform >= 0 & uniform <= (1/6)] <- 1
  dice[uniform > (1/6) & uniform <= (2/6)] <- 2
  dice[uniform > (2/6) & uniform <= (3/6)] <- 3
  dice[uniform > (3/6) & uniform <= (4/6)] <- 4
  dice[uniform > (4/6) & uniform <= (5/6)] <- 5
  dice[uniform > (5/6) & uniform <= (6/6)] <- 6
  
  uniqv <- unique(dice)
  maxnum <- uniqv[which.max(tabulate(match(dice, uniqv)))]
  
  maxnumber[s] <- sum(dice == maxnum)
}

maxnumberdf <- as.data.frame(maxnumber)
p <- ggplot(data=maxnumberdf, aes(x=V1)) +         
  geom_bar( color="black") + 
  scale_x_discrete(limits=c(1:numdice)) + 
  labs(title="1,000,000 sims.of 4 dice",  
       x="Max. of Same Number", 
       y = "Number of Realizations")

p + scale_y_continuous(labels = comma)
```


```{r include=TRUE}
table(maxnumberdf)
```

```{r include=TRUE}
mean(maxnumberdf == 1) * 100

```

#b

```{r include=TRUE}

numdice = 6

numsims = 1000000

maxnumber <- matrix(NA, nrow = numsims, ncol = 1)  

set.seed(12221979)

for (s in 1:numsims) {
  
  uniform <- runif(n = numdice, min = 0, max = 1)

  dice <- matrix(-9, nrow = numdice, ncol = 1)  

  
  dice[uniform >= 0 & uniform <= (1/6)] <- 1
  dice[uniform > (1/6) & uniform <= (2/6)] <- 2
  dice[uniform > (2/6) & uniform <= (3/6)] <- 3
  dice[uniform > (3/6) & uniform <= (4/6)] <- 4
  dice[uniform > (4/6) & uniform <= (5/6)] <- 5
  dice[uniform > (5/6) & uniform <= (6/6)] <- 6
  

  uniqv <- unique(dice)
  maxnum <- uniqv[which.max(tabulate(match(dice, uniqv)))]
  

  maxnumber[s] <- sum(dice == maxnum)
}


maxnumberdf <- as.data.frame(maxnumber)
p <- ggplot(data=maxnumberdf, aes(x=V1)) +         
  geom_bar( color="black") + 
  scale_x_discrete(limits=c(1:numdice)) + 
  labs(title="1,000,000 sims.of 6 dice",  
       x="Max. of Same Number", 
       y = "Number of Realizations")

p + scale_y_continuous(labels = comma)
```

```{r include=TRUE}
table(maxnumberdf)

```

```{r include=TRUE}
mean(maxnumberdf >= 3) * 100

```

#c
```{r include=TRUE}

numdice = 5

numsims = 1000000

maxnumber <- matrix(NA, nrow = numsims, ncol = 1)  

sumdice <- matrix(NA, nrow = numsims, ncol = 1) 

set.seed(12221979)

for (s in 1:numsims) {
  

  uniform <- runif(n = numdice, min = 0, max = 1)

  dice <- matrix(-9, nrow = numdice, ncol = 1)  

  dice[uniform >= 0 & uniform <= (1/6)] <- 1
  dice[uniform > (1/6) & uniform <= (2/6)] <- 2
  dice[uniform > (2/6) & uniform <= (3/6)] <- 3
  dice[uniform > (3/6) & uniform <= (4/6)] <- 4
  dice[uniform > (4/6) & uniform <= (5/6)] <- 5
  dice[uniform > (5/6) & uniform <= (6/6)] <- 6
  

  uniqv <- unique(dice)
  maxnum <- uniqv[which.max(tabulate(match(dice, uniqv)))]
  
  maxnumber[s] <- sum(dice == maxnum)
  
  sumdice[s] <- sum(dice)
}

maxnumberdf <- as.data.frame(maxnumber)
p <- ggplot(data=maxnumberdf, aes(x=V1)) +         
  geom_bar( color="black") + 
  scale_x_discrete(limits=c(1:numdice)) + 
  labs(title="1,000,000 sims.of 5 dice",  
       x="Max. of Same Number", 
       y = "Number of Realizations")

p + scale_y_continuous(labels = comma)
```


```{r include=TRUE}
table(maxnumber)
```
```{r include=TRUE}

table(sumdice)

```

```{r include=TRUE}
100 * mean(maxnumber == 1 & (sumdice == 15 | sumdice == 20))

```

#Question 5
#a

THTHTHHTTTHTHTHTHHTTHTHT

#b
```{r include=TRUE}

set.seed(12221979)


numcoins = 25

numsims = 100000

probheads = 0.5


longstreak <- matrix(NA, nrow = numsims, ncol = 1) 


for (s in 1:numsims) {
  
   
  
  coins <- rbinom(numcoins, size = 1, prob = probheads)
  
  streak <- matrix(NA, nrow = numcoins, ncol = 1) 

  streak[1] <- 1

  for (c in 2:numcoins) {

    streak[c] <- 1 + streak[c-1]*(coins[c] == coins[c-1])

  }
  
  
  longstreak[s] <- max(streak)

}


pdf <- table(longstreak) / numsims

cdf <- pdf
for (c in 2:dim(cdf)) {
  cdf[c] <- cdf[c-1] + pdf[c]
}


cdfdf <- as.data.frame(cdf)

p <- ggplot(data=cdfdf, aes(x=longstreak, y=Freq, group=1)) +
  geom_line()+
  geom_point()+
  labs(title = "CDF in 100,000 Sims.", x = "Length of Streak of Same Coin", 
       y = "Pr(Longest Streak <= Length)")
p

```

#c
```{r include=TRUE}
head(cdfdf)
```

```{r include=TRUE}
(1 - cdfdf$Freq[cdfdf$longstreak == 5]) * 100

```

#d
```{r include=TRUE}
(1 - cdfdf$Freq[cdfdf$longstreak == 3]) * 100

```
l